An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-03-21 EconGrowth-R-Functions.Rnw
Now creating a 3-D graph of SSE vs. lambda and gamma for SF US Q.

  return(graph)
}
@

>>>>

  return(graph)
}

#########################
# Creates a 3-D wireframe graph with SSE on the vertical axis and 
# lambda and m on the horizontal axes.
# params: 
#   countryAbbrev: 2-letter country code (upper case)
#   factor: the factor of production you want to use in the single-factor model.
##
sf3DSSEGraph <- function(countryAbbrev, factor){
  # Generate the data in the form of a matrix with gamma in columns, lambda in rows, and SSE in the middle
  # First, make a data.frame with which we want to make predictions
  lambdaSeq <- seq(0.0, 0.05, by=0.01)
  mSeq <- seq(0.0, 1.0, by=0.1)

  lambdaSeq <- data.frame(lambdaSeq)
  nRowLambda <- nrow(lambdaSeq)
  colnames(lambdaSeq) <- "lambda"

  mSeq <- data.frame(mSeq)
  nRowM <- nrow(mSeq)
  colnames(mSeq) <- "m"

  # Make a sequence of lambda values based on the number of m values you have.
  lambda <- rep(lambdaSeq[1,1], nRowM)
  lambda <- data.frame(lambda)
  colnames(lambda) <- "lambda"
  for (i in 2:nRowLambda){
    lambdaTemp <- rep(lambdaSeq[i, 1], nRowM)
    lambdaTemp <- data.frame(lambdaTemp)
    colnames(lambdaTemp) <- "lambda"
    lambda <- rbind(lambda, lambdaTemp)
  }

  # Make repeats of the m data frame as we have number of lambda rows
  m <- mSeq
  for (i in 2:nRowLambda){
    m <- rbind(m, mSeq)
  }

  # Now create the data.frame of new observation locations
  newX <- cbind(lambda, m)

  # Get the actual data
  data <- loadData(countryAbbrev)
  y_act <- data[ ,"iGDP"] # Pick off the GDP column
  y_act <- data.frame(y_act)

  # Make predictions
  # Get the model
  model <- singleFactorModel(countryAbbrev, factor)

  sse <- c(1:(nRowLambda*nRowM)) # Build a bogus vector at the start

  # Loop over all the different combinations of lambda and m that are of interest for us.
  #i = 1
  for (i in 1:(nRowLambda*nRowM)){
    y_pred <- predict(model, newX[i,]) # Get lambda and m for this particular row.
    y_pred <- data.frame(y_pred)

    compareDF <- cbind(y_pred, y_act)
    se <- with(compareDF, (y_pred-y_act)^2) # Makes a column of squared errors at each year
    compareDF <- cbind(compareDF, se) # Bind it to the data frame as an additional column

    # Now get the sse
    sums <- colSums(compareDF)
    sse[i] <- sums[3] # squared error (se) is the 3rd column
  }
  ssedf <- data.frame(sse)
  colnames(ssedf) <- "sse"

  # Now merge the sse column and the newX data to form a data.frame fit for 3-D plotting
  dataFor3DGraph <- cbind(newX, ssedf)

  # Create the lattice graph
  fig <- wireframe(sse ~ lambda * m,
                   data = dataFor3DGraph,
                   xlab = "$\\lambda$",
                   ylab = "$\\gamma$",
                   zlab = "$SSE$")
  return(fig)    
}
@

### 2013-03-21 EconGrowthSeminar.Rnw
Added new slides to better explain the curve-fit process as optimization.
Now creating a 3-D graph of SSE vs. lambda and gamma for SF US Q.

\begin{frame} %----------------------------
  \begin{center} \huge
    ANALYSIS OF\\
    \vspace{\baselineskip}
    GROWTH MODELS
  \normalsize \end{center}
\end{frame}

@@

<<Coefs_US_SF>>=
sfUSK <- coef(singleFactorModel("US", "K"))
sfUSL <- coef(singleFactorModel("US", "L"))
sfUSQ <- coef(singleFactorModel("US", "Q"))
sfUSX <- coef(singleFactorModel("US", "X"))
sfUSU <- coef(singleFactorModel("US", "U"))
@
\begin{frame}{Single-Factor $y$ Predictions} %----------------------------
  \begin{columns}[c]
    \column{2.2in}
      \begin{center}
        \sfEqn
        \vspace{\baselineskip}

>>>>

\begin{frame} %----------------------------
  \begin{center} \huge
    ANALYSIS OF
    \vspace{\baselineskip}
    \vspace{\baselineskip}
    GROWTH MODELS
  \normalsize \end{center}
\end{frame}

@@

\begin{frame}{Curve-fitting As Optimization} %----------------------------
  \begin{itemize}
    \item{Select values of parameters to minimize errors ($SSE$) between predicted and historical economic output $y$}
    \vspace{\baselineskip}

    \item{Example: Single-factor Model with $q$}
    \begin{itemize}
      \item{$y = \mathrm{e}^{\lambda (t-t_0)}q^{\gamma}$}
      \item{Select $\lambda$ and $\gamma$ to minimize\\
            $SSE = \sum\limits_{t=1980}^{2011}(y_{pred} - y_{act})^2$}
    \end{itemize}
  \end{itemize}
\end{frame}

<<Coefs_US_SF, eval=TRUE>>=
sfUSK <- coef(singleFactorModel("US", "K"))
sfUSL <- coef(singleFactorModel("US", "L"))
sfUSQ <- coef(singleFactorModel("US", "Q"))
sfUSX <- coef(singleFactorModel("US", "X"))
sfUSU <- coef(singleFactorModel("US", "U"))
@

\begin{frame}{$SSE$ Minimization} %---------------------------- 
<<3D_Graph_US_SF_Q, fig.width=presentationGraph1ColWidth, fig.height=presentationGraph1RowHeight, eval=TRUE>>=
sf3DSSEGraph("US", "Q")
@
  \begin{center}
    $y = \mathrm{e}^{\lambda (t-t_0)}q^{\gamma}$
  \end{center}
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putFlagUpperRight{us}{0.055}  
\end{frame}

\begin{frame}{Single-Factor $y$ Predictions} %----------------------------
  \begin{columns}[c]
    \column{2.2in}
      \begin{center}
        \sfEqn
        \vspace{\baselineskip}
